---
title: "Current Conditions"
date: "Last Update: `r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(sf)
library(tidyverse)
library(tmap)
library(rpgcolorsr)
library(rmapshaper)
library(scales)
library(mapdeck)

tmap_mode("view")
far_pal <- rpg_color_pal("rpg_orange_ramp")(10)
key <- paste0("pk.eyJ1IjoiYmZyb2ViLXJwZ3JvdXAiLCJhIjoiY2toY",
              "jM2ODUwMDN2bDJ5cGVpZjU4eWYzbSJ9.",
              "-Fm5keiE7dSTQXK45mI67A")
```

```{r data imports}
parcels  <- read_sf("./data/parcels.shp")
stations <- read_sf("./data/Stations.shp") 
stn_area <- read_sf("./data/StationAreas.shp") 
```

## Known Development
This section shows the details of current and on-going development in the LCRT
corridor. The development forecast depends largely on where development is
already located ("Existing") and where ongoing development activity and/or 
development master plans suggest development will occur in the near term 
("Pipeline"). Use the tabs below to explore floor area by land use in each 
market segment and the density of development associated with the existing
and pipeline phases. The third tab shows the results of adding pipeline 
development to existing development.

## {.tabset}

### Exisiting
Existing development is estimated through the following steps:
  - *Parcel-based* estimates rely on local property appraiser data to determine
  each parcel's land use. Building area on each parcel is estimated using
  building footprint features from Microsoft. <insert link>
  
    - Each building footprint feature is associated with a single parcel based
    on the building's centroid location.
    
    - Total building floor area is estimated based on footprint area, the
    building's estimated height, and a floor height assumption. Some constraints
    apply for different land uses.
  
  - *Point-based* estimates allow known recent developments to be included in
  the existing conditions assessment. Each point represents a recently-delivered
  building and includes attributes showing estimated building square footage and
  land use type. The newness of these developments means they may not be
  reflected in parcel features.
  
The point-based data are used to override parcel-based data for parcels where
recent development activity is observed. Each parcel's land use (from parcel
data or recent development points) is generalized into one of seven classes,
used throughout the forecasting analysis:
  - Single family residential
  - Multi-family residential
  - Hotel
  - Industrial
  - Office
  - Retail
  - Other
  
The chart below shows existing development floor area by land use for each
market segment. 

```{r Existing Sq Ft by LU, out.width= "100%"}
#melt _Ex columns into long form then bar chart, include corridor and segment charts

parcels %>%
  subset(!is.na(seg_nam)) %>%
  select(seg_nam, ends_with("SF_Ex")) %>%
  st_drop_geometry() %>%
  pivot_longer(names_to = "land_use",
               cols = ends_with("Ex"),
               values_to = "square_ft") %>%
  mutate(land_use = str_sub(land_use, 1, -4)) %>%
  mutate(seg_nam  = factor(seg_nam, levels = c("Historic Peninsula", "The Neck",
                                               "South End", "I-526 Employment", 
                                               "Northwoods Mall Rivers Ave", 
                                               "US 78", "Summerville"))) %>%
  ggplot() +
  geom_col(aes(x = seg_nam, y = square_ft, fill = land_use)) +
  scale_fill_rpg(palette = "main and additional",
                 name    = "Land Use",
                 labels  = c("Hotel", "Industrial",
                             "Multi Family", "Office",
                             "Other", "Retail", "Single Family")) +
  scale_y_continuous(labels = comma) +
  ylab("Square Feet") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1),
        axis.title.x = element_blank())
```

The map below shows each parcel's floor area ratio (FAR), which is the total
estimated building area on the parcel divided by its land area. 

```{r Exisiting FAR Map, out.width= "100%"}
# attempting mapdeck
# parcels %>%
#   mutate(bld_area = SF_SF_Ex + MF_SF_Ex + Ret_SF_Ex + Ind_SF_Ex +
#                     Off_SF_Ex + Hot_SF_Ex + Oth_SF_Ex,
#          FAR      = bld_area / Sq_Feet)  %>%
#   fasterize(raster = raster, field = "FAR") %>%
#     tm_shape() + tm_polygon(col = "FAR",
#                              breaks = c(0, 0.25, 0.5, 1, 2, 900),
#                              labels = c("0 to 0.25", "0.25 to 0.50",
#                                         "0.50 to 1.00", "1.00 to 2.00",
#                                         "2.01 or more"),
#                              palette = far_pal,
#                            title = "Floor Area Ratio") +
#   tm_shape(stn_area) + tm_polygons(col = "#D3D3D3", alpha = 0.1) +
#   tm_shape(stations) + tm_dots()
mapdeck(token = key, 
        style = mapdeck_style("dark"),
        location = c(32.8, -79.9)) %>%
  add_sf(
    data = parcels %>%
      mutate(bld_area = SF_SF_Ex + MF_SF_Ex + Ret_SF_Ex + Ind_SF_Ex +
               Off_SF_Ex + Hot_SF_Ex + Oth_SF_Ex,
             FAR      = bld_area / Sq_Feet,
             FAR_Lab  = cut(FAR,
                            breaks = c(0, 0.25, 0.5, 1, 2, 900),
                            labels = c("[0-0.25)", "[0.25-0.5)",
                                       "[0.5-1", "[1-2)", "[2-900)"),
                            include.lowest = TRUE)) %>%
      st_transform(crs = 4326)
    , layer = "polygon_layer"
    , fill_colour = "FAR_Lab"
    , palette = "orrd"
    , legend = TRUE
    , legend_options = list(title = "Floor Area Ratio")
  ) %>%
  add_sf(
    data = stn_area %>% st_transform(crs = 4326)
    , layer = "station_areas"
    #, fill_colour ="#FFFFFF"
    , fill_opacity  = 0
    , stroke_colour = "#FFFFFF"
    , stroke_width  = 15
  ) %>%
  add_sf(
    data = stations %>% st_transform(crs = 4326)
    , layer = "stations"
    , fill_colour = "#D3D3D3"
    , fill_opacity = 255
    , radius_min_pixels = 8
    , radius_max_pixels = 12
  ) 
```

### Pipeline
The development pipeline consists of ongoing or planned development activity,
provided as point locations, similar to the *point-based* existing activity
described in the "Existing" tab. Each point represents a planned or under
construction building and includes attributes showing estimated building square
footage and land use type. The land uses are generalized into seven categories
used throughout the forecasting process.

The pipeline informs the 2040 forecasting process in two ways. First, it records
expected growth that is part of the development future leading up to 2040. 
Second, these sites are removed from consideration for forecasting since their
development future is accounted for in the pipeline analysis.

The chart below shows existing development floor area by land use for each
market segment.

```{r Pipeline Plot, out.width= "100%"}
#summarize the pipe fields in the same way as the _EX columns
parcels %>% 
  subset(!is.na(seg_nam)) %>%
  select(seg_nam, contains("Pip", ignore.case = FALSE)) %>%
  st_drop_geometry() %>%
  pivot_longer(names_to = "land_use",
               cols = contains("Pip"),
               values_to = "square_ft") %>%
  mutate(land_use = str_replace(land_use, "_Pip.?", "")) %>%
  mutate(seg_nam  = factor(seg_nam, levels = c("Historic Peninsula", "The Neck",
                                             "South End", "I-526 Employment", 
                                             "Northwoods Mall Rivers Ave", 
                                             "US 78", "Summerville"))) %>%
  ggplot() +
  geom_col(aes(x = seg_nam, y = square_ft, fill = land_use)) +
  scale_fill_rpg(palette = "main and additional",
                 name    = "Land Use",
                 labels  = c("Hotel", "Industrial",
                             "Multi Family", "Office", 
                             "Other", "Retail", "Single Family")) +
  scale_y_continuous(labels = comma) +
  ylab("Square Feet") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1),
        axis.title.x = element_blank())
```

The map below shows each parcel's FAR for pipeline development only. This is the
planned or under construction floor area divided by the parcel area.

```{r Pipe FAR Map, out.width= "100%"}
# parcels %>%
#   mutate(bld_area = SF_SF_Pipe + MF_SF_Pipe + Ret_SF_Pip + Ind_SF_Pip +
#                     Off_SF_Pip + Hot_SF_Pip + Oth_SF_Pip,
#          FAR      = bld_area / Sq_Feet) %>%
#   tm_shape() + tm_polygons(col = "FAR",
#                              lwd = 0.1,
#                              breaks = c(0, 0.25, 0.5, 1, 2, 900),
#                              labels = c("0 to 0.25", "0.25 to 0.50", 
#                                         "0.50 to 1.00", "1.00 to 2.00", 
#                                         "2.01 or more"),
#                              palette = far_pal,
#                            title = "Floor Area Ratio") +
#   tm_shape(stn_area) + tm_polygons(col = "#D3D3D3", alpha = 0.1) + 
#   tm_shape(stations) + tm_dots()

mapdeck(token = key, 
        style = mapdeck_style("dark"),
        location = c(32.8, -79.9)) %>%
  add_sf(
    data = parcels %>%
      mutate(bld_area = SF_SF_Pipe + MF_SF_Pipe + Ret_SF_Pip + Ind_SF_Pip +
                        Off_SF_Pip + Hot_SF_Pip + Oth_SF_Pip,
             FAR      = bld_area / Sq_Feet,
             FAR_Lab  = cut(FAR,
                            breaks = c(0, 0.25, 0.5, 1, 2, 900),
                            labels = c("[0-0.25)", "[0.25-0.5)",
                                       "[0.5-1", "[1-2)", "[2-900)"),
                            include.lowest = TRUE)) %>%
      st_transform(crs = 4326)
    , layer = "polygon_layer"
    , fill_colour = "FAR_Lab"
    , palette = "orrd"
    , legend = TRUE
    , legend_options = list(title = "Floor Area Ratio")
  ) %>%
  add_sf(
    data = stn_area %>% st_transform(crs = 4326)
    , layer = "station_areas"
    #, fill_colour ="#FFFFFF"
    , fill_opacity  = 0
    , stroke_colour = "#FFFFFF"
    , stroke_width  = 15
  ) %>%
  add_sf(
    data = stations %>% st_transform(crs = 4326)
    , layer = "stations"
    , fill_colour = "#D3D3D3"
    , fill_opacity = 255
    , radius_min_pixels = 8
    , radius_max_pixels = 12
  ) 
```

### Exisiting + Pipeline
This tab shows the results of combining the existing conditions with the 
pipeline data, reflecting all known development. The chart below shows total
known development floor area by land use for each market segment. 

```{r Pipe EX Plot, out.width= "100%"}
#summarize the EX Pipe fields in the same way as the _EX columns
parcels %>% 
  subset(!is.na(seg_nam)) %>%
  select(seg_nam, contains("SF_ExP")) %>%
  st_drop_geometry() %>%
  pivot_longer(names_to = "land_use",
               cols = contains("ExP"),
               values_to = "square_ft") %>%
  mutate(land_use = str_replace(land_use, "_ExP.?", "")) %>%
  mutate(seg_nam  = factor(seg_nam, levels = c("Historic Peninsula", "The Neck",
                                               "South End", "I-526 Employment", 
                                               "Northwoods Mall Rivers Ave", 
                                               "US 78", "Summerville"))) %>%
  ggplot() +
  geom_col(aes(x = seg_nam, y = square_ft, fill = land_use)) +
  scale_fill_rpg(palette = "main and additional",
                 name    = "Land Use",
                 labels  = c("Hotel", "Industrial",
                             "Multi Family", "Office", 
                             "Other", "Retail", "Single Family")) +
  scale_y_continuous(labels = comma) +
  ylab("Square Feet") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1),
        axis.title.x = element_blank())
```

The map below shows FAR estimates for all known development (Existing plus 
pipeline).

```{r Pipe EX FAR Map, out.width= "100%"}
# parcels %>%
#   mutate(bld_area = SF_SF_ExPi + MF_SF_ExPi + Ret_SF_ExP + Ind_SF_ExP +
#                     Off_SF_ExP + Hot_SF_ExP + Oth_SF_ExP,
#          FAR      = bld_area / Sq_Feet) %>%
#   tm_shape() + tm_polygons(col = "FAR",
#                              lwd = 0.1,
#                              breaks = c(0, 0.25, 0.5, 1, 2, 900),
#                              labels = c("0 to 0.25", "0.25 to 0.50", 
#                                         "0.50 to 1.00", "1.00 to 2.00", 
#                                         "2.01 or more"),
#                              palette = far_pal,
#                            title = "Floor Area Ratio") +
#   tm_shape(stn_area) + tm_polygons(col = "#D3D3D3", alpha = 0.1) + 
#   tm_shape(stations) + tm_dots()

mapdeck(token = key, 
        style = mapdeck_style("dark"),
        location = c(32.8, -79.9)) %>%
  add_sf(
    data = parcels %>%
      mutate(bld_area = SF_SF_ExPi + MF_SF_ExPi + Ret_SF_ExP + Ind_SF_ExP +
                        Off_SF_ExP + Hot_SF_ExP + Oth_SF_ExP,
             FAR      = bld_area / Sq_Feet,
             FAR_Lab  = cut(FAR,
                            breaks = c(0, 0.25, 0.5, 1, 2, 900),
                            labels = c("[0-0.25)", "[0.25-0.5)",
                                       "[0.5-1", "[1-2)", "[2-900)"),
                            include.lowest = TRUE)) %>%
      st_transform(crs = 4326)
    , layer = "polygon_layer"
    , fill_colour = "FAR_Lab"
    , palette = "orrd"
    , legend = TRUE
    , legend_options = list(title = "Floor Area Ratio")
  ) %>%
  add_sf(
    data = stn_area %>% st_transform(crs = 4326)
    , layer = "station_areas"
    #, fill_colour ="#FFFFFF"
    , fill_opacity  = 0
    , stroke_colour = "#FFFFFF"
    , stroke_width  = 15
  ) %>%
  add_sf(
    data = stations %>% st_transform(crs = 4326)
    , layer = "stations"
    , fill_colour = "#D3D3D3"
    , fill_opacity = 255
    , radius_min_pixels = 8
    , radius_max_pixels = 12
  ) 
```