---
title: "Current Conditions"
date: "Last Update: `r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(sf)
library(tidyverse)
library(tmap)
library(rpgcolorsr)
library(rmapshaper)
library(scales)
library(mapdeck)

tmap_mode("view")
far_pal <- rpg_color_pal("rpg_orange_ramp")(10)
key <- paste0("pk.eyJ1IjoiYmZyb2ViLXJwZ3JvdXAiLCJhIjoiY2toY",
              "jM2ODUwMDN2bDJ5cGVpZjU4eWYzbSJ9.",
              "-Fm5keiE7dSTQXK45mI67A")
```

```{r data imports}
segments  <- read_sf("./data/corridor_marketsegments.shp")
stations    <- read_sf("./data/Stations.shp") 
stn_area    <- read_sf("./data/StationAreas.shp") 
```

This site provides an overview of the land use forecasting process used to
develop estimates of jobs, residents, and building floor area in the Low
Country Rapid Transit (LCRT) corridor from the Downtown Charleston peninsula
to Summerville. 

The objective of the forecasting process is to update 2040 socioeconomic 
estimates that inform regional travel models and assessments the ridership
potential of BRT.

To account for the influence of transit stations on local development potential,
the forecasting process focuses on individual parcels and results are 
summarized to the travel analysis zones (TAZs) used in ridership modeling. Both
parcels and TAZs are orgnized into seven major segments, each representing
distinct development markets in the LCRT corridor.

The forecasting process is organized into three key phases:

  - **Known development**: How much development is currently on the ground?
  Where is development activity ongoing? 
  
  - **Allocation preparation**: How much new development is reasonable to expect
  by 2040? What sites are suitable for development? What kind and amount of 
  development is appropriate at each site?
  
  - **Allocation results**: Where is growth through 2040 expected to go,
  assuming BRT station locations and typologies?


The map below shows the LCRT corridor with its potential transit station 
locations and the seven market segments used in the foreacsting analysis.

```{r Station and Segment Map, , out.width= "100%"}
tm_basemap(paste0("http://services.arcgisonline.com/arcgis/rest/services/Canvas/",
    "World_Dark_Gray_Base/MapServer/tile/{z}/{y}/{x}")) +
  tm_shape(segments) + tm_polygons(col = "SegmentNm",
                                   title = "Segment") +
  tm_shape(stn_area) + tm_polygons(col = "#D3D3D3", alpha = 0.1) +
  tm_shape(stations) + tm_dots() + tm_text("stn_name", just = "bottom", size = 0.75) +
  tm_tiles(paste0("http://services.arcgisonline.com/arcgis/rest/services/Canvas/",
    "World_Light_Gray_Reference/MapServer/tile/{z}/{y}/{x}"), group = "Labels")

```

