---
title: "Allocation Prep"
author: "Brian Froeb"
date: "11/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(sf)
library(tidyverse)
library(tmap)
library(rpgcolorsr)
library(rmapshaper)
library(scales)
library(mapdeck)

tmap_mode("view")
far_pal <- rpg_color_pal("rpg_orange_ramp")(10)
cap_pal <- rpg_color_pal("rpg_blue_ramp")(10)
suit_pal <- rpg_color_pal("rpg_colors")(10)
key <- paste0("pk.eyJ1IjoiYmZyb2ViLXJwZ3JvdXAiLCJhIjoiY2toY",
              "jM2ODUwMDN2bDJ5cGVpZjU4eWYzbSJ9.",
              "-Fm5keiE7dSTQXK45mI67A")
```

```{r}
parcels  <- read_sf("./data/parcels.shp")
stations <- read_sf("./data/Stations.shp") 
stn_area <- read_sf("./data/StationAreas.shp") 
```
## Processing Steps

## {.tabset}

### Suitability

Parcels can have zero suitability because:

- They are single family homes and not likely to change

- They are in the pipeline and have already been accounted for
-- This includes parcels that are apart of master planned developments

Parcels that have been identified as a "development opportunity sites" are rated as the highest suitability.

```{r Suit Map, echo=FALSE, eval=FALSE}
# Zero_Suit   <- subset(parcels, parcels$tod_suit == 0)
# Suitability <- subset(parcels, parcels$tod_suit != 0)
# 
# Dev_Sites <- subset(parcels, parcels$DO_Site == 1)
# Dev_Sites$PrtlDOS <- as.factor(Dev_Sites$PrtlDOS)
# 
# # multiple layer tmap map, Dev Sites (Partial navy #40516f Full purple #8400a8), 0 suit gray, suit ramp blue to teal
# # add station and station areas
# 
# tm_shape(Zero_Suit) + tm_polygons(col = "#D3D3D3",
#                                   lwd = 0.1,
#                                   labels = "No Suitability") +
#   tm_shape(Suitability) + tm_polygons(col = "tod_suit",
#                                       lwd = 0.1,
#                                       palette = suit_pal,
#                                       title = "Suitability") +
#   tm_shape(Dev_Sites) + tm_polygons(col = "PrtlDOS",
#                                     lwd = 0.1,
#                                     palette = c("#8400a8", "#40516f"),
#                                     labels = c("Full DO Site", "Partial DO Site"),
#                                     title = "Developmet Sites") +
#   tm_shape(stn_area) + tm_polygons(col = "#D3D3D3", alpha = 0.1) + 
#   tm_shape(stations) + tm_dots()
```

Because mapdeck does not allow for turning layers on and off I elevated Partial DO Sites so that that dimension can be retained. Let me know what you think.

```{r}
mapdeck(token = key, 
        style = mapdeck_style("dark"),
        location = c(32.8, -79.9)) %>%
  add_sf(
      data = parcels %>%
        mutate(PrtlDOS = PrtlDOS * 100) %>%
        st_transform(crs = 4326)
    , layer = "polygon_layer"
    , fill_colour = "tod_suit"
    , palette = "viridis"
    , legend = TRUE
    , legend_options = list(title = "TOD Suitability")
    , elevation = "PrtlDOS"
  ) %>%
  add_sf(
    data = stn_area %>% st_transform(crs = 4326)
    , layer = "station_areas"
    #, fill_colour ="#FFFFFF"
    , fill_opacity  = 0
    , stroke_colour = "#FFFFFF"
    , stroke_width  = 15
  ) %>%
  add_sf(
    data = stations %>% st_transform(crs = 4326)
    , layer = "stations"
    , fill_colour = "#D3D3D3"
    , fill_opacity = 255
    , radius_min_pixels = 8
    , radius_max_pixels = 12
  ) 
```

### Base Line Capacity

```{r}
# ggplot of bcap and seg, just like first page
parcels %>% 
  select(seg_nam, contains("BCa")) %>%
  st_drop_geometry() %>%
  pivot_longer(names_to = "land_use",
               cols = contains("BCa"),
               values_to = "square_ft") %>%
  mutate(land_use = str_replace(land_use, "_BCa.?", "")) %>%
  ggplot() +
  geom_col(aes(x = seg_nam, y = square_ft, fill = land_use)) +
  scale_fill_rpg(palette = "main and additional",
                 name    = "Land Use",
                 labels  = c("Hotel", "Industrial",
                             "Multi Family", "Office", 
                             "Other", "Retail", "Single Family")) +
  scale_y_continuous(labels = comma) +
  ylab("Square Feet") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1),
        axis.title.x = element_blank())
```


```{r}
# Mean Far Map
# parcels %>%
#     tm_shape() + tm_polygons(col = "Mean_FAR",
#                              lwd = 0.1,
#                              breaks = c(0, 0.25, 0.5, 1, 2, 900),
#                              labels = c("0 to 0.25", "0.25 to 0.50", 
#                                         "0.50 to 1.00", "1.00 to 2.00", 
#                                         "2.01 or more"),
#                              palette = far_pal,
#                              title = "Floor Area Ratio") +
#   tm_shape(stn_area) + tm_polygons(col = "#D3D3D3", alpha = 0.1) + 
#   tm_shape(stations) + tm_dots()

mapdeck(token = key, 
        style = mapdeck_style("dark"),
        location = c(32.8, -79.9)) %>%
  add_sf(
    data = parcels %>%
      mutate(FAR_Lab  = cut(Mean_FAR,
                            breaks = c(0, 0.25, 0.5, 1, 2, 900),
                            labels = c("[0-0.25)", "[0.25-0.5)",
                                       "[0.5-1", "[1-2)", "[2-900)"),
                            include.lowest = TRUE)) %>%
      st_transform(crs = 4326)
    , layer = "polygon_layer"
    , fill_colour = "FAR_Lab"
    , palette = "orrd"
    , legend = TRUE
    , legend_options = list(title = "Floor Area Ratio")
  ) %>%
  add_sf(
    data = stn_area %>% st_transform(crs = 4326)
    , layer = "station_areas"
    #, fill_colour ="#FFFFFF"
    , fill_opacity  = 0
    , stroke_colour = "#FFFFFF"
    , stroke_width  = 15
  ) %>%
  add_sf(
    data = stations %>% st_transform(crs = 4326)
    , layer = "stations"
    , fill_colour = "#D3D3D3"
    , fill_opacity = 255
    , radius_min_pixels = 8
    , radius_max_pixels = 12
  ) 
```

This is the baseline for the allocation process out side of stations areas, parcels within station areas receive additional processing.

### Build Out Capacity

- Each station area, at full build out, will have enough capacity to meet certain targets set by the station typology. 

- We assume that development is likely to be most intense near to the station and on the high suitability sites. 

- We assume that residential development will become a higher proportion of total development at the edges of the station area. 

- Existing development, including development with no suitability for change, counts towards the activity targets and influences the distribution of additional growth within a station area. 

- Pipeline and Existing development are accounted for (changed wording)

<br />

```{r TOD Cap Plot}
parcels %>% 
  select(seg_nam, contains("SF_Tot", ignore.case = FALSE)) %>%
  st_drop_geometry() %>%
  mutate(Oth_SF_Tot = 0) %>%
   pivot_longer(names_to = "land_use",
               cols = contains("Tot"),
               values_to = "square_ft") %>%
  filter(!is.null(land_use)) %>%
  filter(!is.na(seg_nam)) %>%
  mutate(land_use = str_replace(land_use, "Tot.?", "")) %>%
  ggplot() +
  geom_col(aes(x = seg_nam, y = square_ft, fill = land_use)) +
  scale_fill_rpg(palette = "main and additional",
                 name    = "Land Use",
                 labels  = c("Hotel", "Industrial",
                             "Multi Family", "Office", 
                             "Other", "Retail", "Single Family")) +
  scale_y_continuous(labels = comma) +
  ylab("Square Feet") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1),
        axis.title.x = element_blank())
```

```{r Tot map}
#plot parcel total cap after summing the tot_cap columns also break down by segment
# parcels %>%
#     mutate(Capacity = SF_SF_TotC + MF_SF_TotC + Ret_SF_Tot + 
#                       Ind_SF_Tot + Off_SF_Tot + Hot_SF_Tot) %>%
#     tm_shape() + tm_polygons(col = "Capacity",
#                              lwd = 0.1,
#                              palette = cap_pal,
#                              title = "Floor Area") +
#   tm_shape(stn_area) + tm_polygons(col = "#D3D3D3", alpha = 0.1) + 
#   tm_shape(stations) + tm_dots()

mapdeck(token = key, 
        style = mapdeck_style("dark")) %>%
  add_sf(
    data = parcels %>%
      mutate(Capacity = SF_SF_TotC + MF_SF_TotC + Ret_SF_Tot + 
                        Ind_SF_Tot + Off_SF_Tot + Hot_SF_Tot,
             FAR      = Capacity / Sq_Feet,
             FAR_Lab  = cut(FAR,
                            breaks = c(0, 0.25, 0.5, 1, 2, 900),
                            labels = c("[0-0.25)", "[0.25-0.5)",
                                       "[0.5-1", "[1-2)", "[2-900)"),
                            include.lowest = TRUE)) %>%
      st_transform(crs = 4326)
    , layer = "polygon_layer"
    , fill_colour = "FAR_Lab"
    , palette = "orrd"
    , legend = TRUE
    , legend_options = list(title = "Capacity Floor Area Ratio")
  ) %>%
  add_sf(
    data = stn_area %>% st_transform(crs = 4326)
    , layer = "station_areas"
    #, fill_colour ="#FFFFFF"
    , fill_opacity  = 0
    , stroke_colour = "#FFFFFF"
    , stroke_width  = 15
  ) %>%
  add_sf(
    data = stations %>% st_transform(crs = 4326)
    , layer = "stations"
    , fill_colour = "#D3D3D3"
    , fill_opacity = 255
    , radius_min_pixels = 8
    , radius_max_pixels = 12
  ) 
```

Add a transition paragraph