---
title: "Allocation Preparation"
date: "Last Update: `r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(sf)
library(tidyverse)
library(tmap)
library(rpgcolorsr)
library(rmapshaper)
library(scales)
library(mapdeck)
library(kableExtra)

tmap_mode("view")
far_pal <- rpg_color_pal("rpg_orange_ramp")(10)
cap_pal <- rpg_color_pal("rpg_blue_ramp")(10)
suit_pal <- rpg_color_pal("rpg_colors")(10)
key <- paste0("pk.eyJ1IjoiYmZyb2ViLXJwZ3JvdXAiLCJhIjoiY2toY",
              "jM2ODUwMDN2bDJ5cGVpZjU4eWYzbSJ9.",
              "-Fm5keiE7dSTQXK45mI67A")
```

```{r Imports}
parcels  <- read_sf("./data/parcels.shp")
stations <- read_sf("./data/Stations.shp") 
stn_area <- read_sf("./data/StationAreas.shp") 
types    <- read_csv("./data/Station_types.csv")
```

## Overview

The 2040 parcel-level land use forecast entails allocating expected increments
of floor area by land use type to appropriate development sites. This requires
assessing each parcel's suitability for development and estimating a plausible
development program for each site. During allocation, the most suitable sites
are filled first, each according to its estimated development program, until
all total expected floor area is distributed among eligible sites.

This section presents the components and results of the development suitability
assessment and the estimation of each parcel's potential development program, 
also referred to as "capacity". Capacity estimation is broken out in two phases:
*baseline capacity* is a simple assessment of development potential based on
site characteristics and prevailing development contexts; *build out capacity*
adjusts baseline estimates for sites in TOD station areas based on station
typologies and associated development targets.

## {.tabset}

### Suitability
Parcel development suitability is estimated based on five criteria:

1. The parcel has been flagged as a "development opportunity site" (DO site). 
This is 60% of the suitability score, scaled by the proportion of the site in
the DO site area (some parcels have partial overlap with DO sites).

2. The parcel is vacant. This is 15% of the suitability score.

3. The parcel is within a TOD station area (half-mile simple buffer around
station locations). This is 5% of the suitability score.

4. The parcel is within the TOD walkshed (quarter-mile network-based buffer
around station  locations). This is 10% of the suitability score.

5. Site size. Larger sites are more suitable than smaller sites. This is 10%
of the suitability score.

Parcels may have no suitability because:

- They are single family homes and not likely to change

- They are in the pipeline and have already been accounted for (this includes 
parcels that are apart of master planned developments)

The map below shows the suitability evaluation results. Parcels are extruded based on their suitability score and colored based on the proportion of a parcel that overlaps a development opportunity site (yellow  is a full site).

```{r Suit Map Old, echo=FALSE, eval=FALSE}
# Zero_Suit   <- subset(parcels, parcels$tod_suit == 0)
# Suitability <- subset(parcels, parcels$tod_suit != 0)
# 
# Dev_Sites <- subset(parcels, parcels$DO_Site == 1)
# Dev_Sites$PrtlDOS <- as.factor(Dev_Sites$PrtlDOS)
# 
# # multiple layer tmap map, Dev Sites (Partial navy #40516f Full purple #8400a8), 0 suit gray, suit ramp blue to teal
# # add station and station areas
# 
# tm_shape(Zero_Suit) + tm_polygons(col = "#D3D3D3",
#                                   lwd = 0.1,
#                                   labels = "No Suitability") +
#   tm_shape(Suitability) + tm_polygons(col = "tod_suit",
#                                       lwd = 0.1,
#                                       palette = suit_pal,
#                                       title = "Suitability") +
#   tm_shape(Dev_Sites) + tm_polygons(col = "PrtlDOS",
#                                     lwd = 0.1,
#                                     palette = c("#8400a8", "#40516f"),
#                                     labels = c("Full DO Site", "Partial DO Site"),
#                                     title = "Developmet Sites") +
#   tm_shape(stn_area) + tm_polygons(col = "#D3D3D3", alpha = 0.1) + 
#   tm_shape(stations) + tm_dots()
```



```{r Suit Map, out.width= "100%"}
mapdeck(token = key, 
        style = mapdeck_style("dark"),
        location = c(32.8, -79.9)) %>%
  add_sf(
    data = parcels %>%
      #subset(DO_Site != 0) %>%
      mutate(PrtlDOS  = as.factor(PrtlDOS),
             Area_AC = (DOSProp * 100)) %>%
      st_transform(crs = 4326)
    , layer = "polygon_layer"
    , fill_colour =  "DOSProp" #"PrtlDOS"
    , palette = "viridis"
    , legend = TRUE
    , legend_options = list(title = "DO Site Proportion")
    , elevation = "tod_suit"
    , elevation_scale = 100
  ) %>%
  add_sf(
    data = stn_area %>% st_transform(crs = 4326)
    , layer = "station_areas"
    #, fill_colour ="#FFFFFF"
    , fill_opacity  = 0
    , stroke_colour = "#FFFFFF"
    , stroke_width  = 15
  ) %>%
  add_sf(
    data = stations %>% st_transform(crs = 4326)
    , layer = "stations"
    , fill_colour = "#D3D3D3"
    , fill_opacity = 255
    , radius_min_pixels = 8
    , radius_max_pixels = 12
  ) 
```

### Baseline Capacity
Baseline development capacity is estimated throughout the corridor based on
parcel characteristics and prevailing development patterns.

- Parcels with active uses are assumed to stay the same land use in the future.

- Vacant/undeveloped parcels are assumed to offer development opportunities 
whose land uses are determined as follows:

  - Small parcels (less than a quarter-acre in area) will be single family
  residential.
  
  - Other parcels' land uses will match the most commonly observed land use
  (weighted by floor area) among neighboring parcels (within a quarter mile).
  
- Intensity of development can be as high as the mean FAR observed among 
neighboring parcels whose land use matches the parcel's expected land use or
the current FAR, whichever is greater.

The chart below shows the baseline capacity estimated by land use for each 
market segment.

```{r baseline cap, out.width= "100%"}
# ggplot of bcap and seg, just like first page
parcels %>% 
  select(seg_nam, contains("BCa")) %>%
  st_drop_geometry() %>%
  pivot_longer(names_to = "land_use",
               cols = contains("BCa"),
               values_to = "square_ft") %>%
  mutate(land_use = str_replace(land_use, "_BCa.?", "")) %>%
  mutate(seg_nam  = factor(seg_nam, levels = c("Historic Peninsula", "The Neck",
                                               "South End", "I-526 Employment", 
                                               "Northwoods Mall Rivers Ave", 
                                               "US 78", "Summerville"))) %>%
  ggplot() +
  geom_col(aes(x = seg_nam, y = square_ft, fill = land_use)) +
  scale_fill_rpg(palette = "main and additional",
                 name    = "Land Use",
                 labels  = c("Hotel", "Industrial",
                             "Multi Family", "Office", 
                             "Other", "Retail", "Single Family")) +
  scale_y_continuous(labels = comma) +
  ylab("Square Feet") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1),
        axis.title.x = element_blank())
```

The map below shows the maximum FAR projected under the baseline capacity
assumptions. Note that while capacity may exist for relatively dense development,
this does not mean that development will be forecasted to attain those densities
in each parcel. It is simply a maximum under the simple assumptions described 
above.

```{r mean far map, out.width= "100%"}
# Mean Far Map
# parcels %>%
#     tm_shape() + tm_polygons(col = "Mean_FAR",
#                              lwd = 0.1,
#                              breaks = c(0, 0.25, 0.5, 1, 2, 900),
#                              labels = c("0 to 0.25", "0.25 to 0.50", 
#                                         "0.50 to 1.00", "1.00 to 2.00", 
#                                         "2.01 or more"),
#                              palette = far_pal,
#                              title = "Floor Area Ratio") +
#   tm_shape(stn_area) + tm_polygons(col = "#D3D3D3", alpha = 0.1) + 
#   tm_shape(stations) + tm_dots()

mapdeck(token = key, 
        style = mapdeck_style("dark"),
        location = c(32.8, -79.9)) %>%
  add_sf(
    data = parcels %>%
      mutate(Capacity = SF_SF_BCap + MF_SF_BCap + Ret_SF_BCa + 
                        Ind_SF_BCa + Off_SF_BCa + Hot_SF_BCa,
             FAR      = Capacity / Sq_Feet,
             FAR_Lab  = cut(FAR,
                             breaks = c(0, 0.25, 0.5, 1, 2, 900),
                             labels = c("[0-0.25)", "[0.25-0.5)",
                                        "[0.5-1", "[1-2)", "[2 and up)"),
                             include.lowest = TRUE)) %>%
      st_transform(crs = 4326)
    , layer = "polygon_layer"
    , fill_colour = "FAR_Lab"
    , palette = "orrd"
    , legend = TRUE
    , legend_options = list(title = "Floor Area Ratio")
  ) %>%
  add_sf(
    data = stn_area %>% st_transform(crs = 4326)
    , layer = "station_areas"
    #, fill_colour ="#FFFFFF"
    , fill_opacity  = 0
    , stroke_colour = "#FFFFFF"
    , stroke_width  = 15
  ) %>%
  add_sf(
    data = stations %>% st_transform(crs = 4326)
    , layer = "stations"
    , fill_colour = "#D3D3D3"
    , fill_opacity = 255
    , radius_min_pixels = 8
    , radius_max_pixels = 12
  ) 
```


### Build Out Capacity
The baseline capacity procedure is run for all parcels throughout the LCRT 
corridor. For sites within TOD station areas, however, a separate capacity 
analysis is undertaken, the results of which overwrite the baseline estimates.
Total build out capacity, then, is a blend of baseline assumptions (outside
station areas) and TOD assumptions (within station areas).

TOD station area capacity assumptions:

- Each station area, at full build out, will have enough capacity to meet 
development targets set by the station typology for that station. 

- Development is likely to be most intense near the station and on high
suitability sites. Development intensities will generally transition downward
for sites at the edges of the station area.

- Residential development represents an increasingly large proportion of total
development at the edges of the station area. Residential development is still
expected near station areas but in lower relative proportions than at the
edges.

- Existing and pipeline development counts towards the attainment of activity
targets and influences the distribution of additional growth within a station 
area.

- Parcels with no suitability for change (single family residential, pipeline
parcels, e.g.) are not subject to capacity estimation, but their estimated
floor areas contribute to station area target attainment as noted above.

The chart below shows the total build out floor area capacity by land use
and market segment.

<br />

```{r TOD Cap Plot, out.width= "100%"}
parcels %>% 
  select(seg_nam, contains("SF_Tot", ignore.case = FALSE)) %>%
  st_drop_geometry() %>%
  mutate(Oth_SF_Tot = 0) %>%
   pivot_longer(names_to = "land_use",
               cols = contains("Tot"),
               values_to = "square_ft") %>%
  filter(!is.null(land_use)) %>%
  filter(!is.na(seg_nam)) %>%
  mutate(land_use = str_replace(land_use, "Tot.?", "")) %>%
  mutate(seg_nam  = factor(seg_nam, levels = c("Historic Peninsula", "The Neck",
                                             "South End", "I-526 Employment", 
                                             "Northwoods Mall Rivers Ave", 
                                             "US 78", "Summerville"))) %>%
  ggplot() +
  geom_col(aes(x = seg_nam, y = square_ft, fill = land_use)) +
  scale_fill_rpg(palette = "main and additional",
                 name    = "Land Use",
                 labels  = c("Hotel", "Industrial",
                             "Multi Family", "Office", 
                             "Other", "Retail", "Single Family")) +
  scale_y_continuous(labels = comma) +
  ylab("Square Feet") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1),
        axis.title.x = element_blank())
```

The map below shows the maximum FAR projected for total build out capacity. It
is identical to baseline capacity outside of station areas and is modified
based on TOD parameters inside station area. As with the baseline capacity map,
the FAR values shown do not imply that development will be forecasted to attain 
those densities in each parcel. It is simply a maximum under the assumptions
described above.

```{r Tot map, out.width= "100%"}
#plot parcel total cap after summing the tot_cap columns also break down by segment
# parcels %>%
#     mutate(Capacity = SF_SF_TotC + MF_SF_TotC + Ret_SF_Tot + 
#                       Ind_SF_Tot + Off_SF_Tot + Hot_SF_Tot) %>%
#     tm_shape() + tm_polygons(col = "Capacity",
#                              lwd = 0.1,
#                              palette = cap_pal,
#                              title = "Floor Area") +
#   tm_shape(stn_area) + tm_polygons(col = "#D3D3D3", alpha = 0.1) + 
#   tm_shape(stations) + tm_dots()

mapdeck(token = key, 
        style = mapdeck_style("dark")) %>%
  add_sf(
    data = parcels %>%
      mutate(Capacity = SF_SF_TotC + MF_SF_TotC + Ret_SF_Tot + 
                        Ind_SF_Tot + Off_SF_Tot + Hot_SF_Tot,
             FAR      = Capacity / Sq_Feet,
             FAR_Lab  = cut(FAR,
                            breaks = c(0, 0.25, 0.5, 1, 2, 900),
                            labels = c("[0-0.25)", "[0.25-0.5)",
                                       "[0.5-1", "[1-2)", "[2 and up)"),
                            include.lowest = TRUE)) %>%
      st_transform(crs = 4326)
    , layer = "polygon_layer"
    , fill_colour = "FAR_Lab"
    , palette = "orrd"
    , legend = TRUE
    , legend_options = list(title = "Capacity Floor Area Ratio")
  ) %>%
  add_sf(
    data = stn_area %>% st_transform(crs = 4326)
    , layer = "station_areas"
    #, fill_colour ="#FFFFFF"
    , fill_opacity  = 0
    , stroke_colour = "#FFFFFF"
    , stroke_width  = 15
  ) %>%
  add_sf(
    data = stations %>% st_transform(crs = 4326)
    , layer = "stations"
    , fill_colour = "#D3D3D3"
    , fill_opacity = 255
    , radius_min_pixels = 8
    , radius_max_pixels = 12
  ) 
```

The table below lists each station with its type and associated targets for
residential units and jobs. Stations are listed in order from north (Summerville)
to south (West Edge)


```{r Job Target, out.width= "100%"}
types %>%
  select(stn_name,stn_type, res_target, job_target) %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

