---
title: "Allocation Results Summary"
date: "Last Update: `r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(sf)
library(tidyverse)
library(tmap)
library(rpgcolorsr)
library(rmapshaper)
library(scales)
library(mapdeck)
library(kableExtra)

tmap_mode("view")
far_pal <- rpg_color_pal("rpg_orange_ramp")(10)
cap_pal <- rpg_color_pal("rpg_blue_ramp")(10)
suit_pal <- rpg_color_pal("rpg_colors")(10)
key <- paste0("pk.eyJ1IjoiYmZyb2ViLXJwZ3JvdXAiLCJhIjoiY2toY",
              "jM2ODUwMDN2bDJ5cGVpZjU4eWYzbSJ9.",
              "-Fm5keiE7dSTQXK45mI67A")
```

```{r Data Import}
stations <- read_sf("./data/Stations.shp") 
stn_area <- read_sf("./data/StationAreas.shp") 
control  <- read_csv("./data/ControlTotals.csv")
segments <- read_sf("./data/corridor_marketsegments.shp")
```

Below is a rundown of the allocation results for pipeline development
and post-pipeline development (allocation). Each section presents bar charts
and tables showing the control totals estimated for each market segment and
the amount of growth modeled through the allocation process.

## {.tabset}

### Pipeline Controls and Values
Pipeline development control totals come from a tabulation of master planned
developments, summarized by segment. These estimates are reflected in the
pipeline points features used to estimate parcel pipeline development. Those
points also included several smaller planned developments not included in the
master planned developments roster. Therefor, the total floor area modeled in
the pipeline exceeds the control totals by about 25 percent.
```{r ggplot of Control Totals, out.width= "100%"}
st_drop_geometry(segments) %>% 
  select(SegmentNm, SegmentNum) %>%
  left_join(control, by = c("SegmentNum" = "Segment")) %>%
  select(SegmentNum, SegmentNm, `Land Use`, PipelineControl, Pipeline) %>%
  arrange(SegmentNum) %>%
  select(SegmentNm, `Land Use`, PipelineControl, Pipeline) %>%
  mutate(PipelineControl = as.numeric(PipelineControl),
         Pipeline         = as.numeric(Pipeline)) %>%
  pivot_longer(contains("pipe"),
               names_to = "control",
               values_to = "sq_feet") %>%
  mutate(SegmentNm  = factor(SegmentNm, levels = c("Historic Peninsula", "The Neck",
                                               "South End", "I-526 Employment", 
                                               "Northwoods Mall Rivers Ave", 
                                               "US 78", "Summerville"))) %>%
  mutate(control = factor(control, levels = c("PipelineControl", "Pipeline"))) %>%
  ggplot() +
  geom_col(aes(x = SegmentNm, y = sq_feet, fill = `Land Use`)) +
  facet_grid(~ control) +
  scale_fill_rpg(palette = "main and additional",
                 name    = "Land Use",
                 labels  = c("Hotel", "Industrial",
                             "Multi Family", "Office", 
                             "Other", "Retail", "Single Family")) +
  scale_y_continuous(labels = comma) +
  ylab("Square Feet") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1),
        axis.title.x = element_blank())
# 
#  ggplot() +
#   geom_col(aes(x = seg_nam, y = square_ft, fill = land_use)) +
#   scale_fill_rpg(palette = "main and additional",
#                  name    = "Land Use",
#                  labels  = c("Hotel", "Industrial",
#                              "Multi Family", "Office", 
#                              "Other", "Retail", "Single Family")) +
#   scale_y_continuous(labels = comma) +
#   ylab("Square Feet") +
#   theme(axis.text.x = element_text(angle = 45,
#                                    hjust = 1),
#         axis.title.x = element_blank())
```

```{r PipelingControl Totals Table, out.width= "100%"}
st_drop_geometry(segments) %>% 
  select(SegmentNm, SegmentNum) %>%
  left_join(control, by = c("SegmentNum" = "Segment")) %>%
  select(SegmentNum, SegmentNm, `Land Use`, PipelineControl, Pipeline) %>%
  arrange(SegmentNum) %>%
  select(SegmentNm, `Land Use`, PipelineControl, Pipeline) %>%
  mutate(PipelineControl = as.numeric(PipelineControl),
         Pipeline        = as.numeric(Pipeline)) %>%
  kable() %>%
  collapse_rows(columns = 1:2, valign = "middle") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 
```



### Allocation Control Totals and Allocated Values

Allocation control totals were tabulated by market segment based on each 
segment's characteristics and development opportunities. With the exception of
a modest number of hotel rooms in the Summerville segment, all forecasted
development was allocated to parecels.

```{r ggplot of allocation Totals, out.width= "100%"}
st_drop_geometry(segments) %>% 
  select(SegmentNm, SegmentNum) %>%
  left_join(control , by = c("SegmentNum" = "Segment")) %>%
  select(SegmentNum, SegmentNm, `Land Use`, AllocationControl, Allocated) %>%
  arrange(SegmentNum) %>%
  select(SegmentNm, `Land Use`, AllocationControl, Allocated) %>%
  mutate(AllocationControl = as.numeric(AllocationControl),
         Allocated         = as.numeric(Allocated)) %>%
  pivot_longer(contains("allo"),
               names_to = "control",
               values_to = "sq_feet") %>%
  mutate(SegmentNm  = factor(SegmentNm, levels = c("Historic Peninsula", "The Neck",
                                             "South End", "I-526 Employment", 
                                             "Northwoods Mall Rivers Ave", 
                                             "US 78", "Summerville"))) %>%
  mutate(control = factor(control, levels = c("AllocationControl", "Allocated"))) %>%
  ggplot() +
  geom_col(aes(x = SegmentNm, y = sq_feet, fill = `Land Use`)) +
  facet_grid(~ control) +
  scale_fill_rpg(palette = "main and additional",
                 name    = "Land Use",
                 labels  = c("Hotel", "Industrial",
                             "Multi Family", "Office", 
                             "Other", "Retail", "Single Family")) +
  scale_y_continuous(labels = comma) +
  ylab("Square Feet") +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1),
        axis.title.x = element_blank())
```


```{r Table of Allocation Totals, out.width= "100%"}
st_drop_geometry(segments) %>% 
  select(SegmentNm, SegmentNum) %>%
  left_join(control, by = c("SegmentNum" = "Segment")) %>%
  select(SegmentNum, SegmentNm, `Land Use`, AllocationControl, Allocated) %>%
  arrange(SegmentNum) %>%
  select(SegmentNm, `Land Use`, AllocationControl, Allocated) %>%
  kable() %>%
  collapse_rows(columns = 1:2, valign = "middle") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```



Full results available [here](https://storymaps.arcgis.com/stories/2d60f4d4609a4d798e165d69ebd46b65)